Convergence Studies and Effective Alphas for Disks with Gravitational Instabilities. \\ I. The Constant Cooling Case

### 2010-10-29 Paper_1.tex
Made changes to eqns 5 & 6
	modified:   Paper_1.tex
	new file:   figures/.gitignore

\usepackage{color}
\usepackage{graphicx}
\usepackage{amsmath, amsthm, amssymb}

@@

Figure 2
%\ref{fig:DensityPlots}, 
shows midplane and radial densities  at $t \sim 18$ orps  
for the $l_{max} =$ 64, 128, 256, and 512 simulations. 
The lower resolution (smaller $l_{max}$) simulations
display longer wavelength spiral structures, with larger amplitudes and more coherence than those
seen in higher resolution 
simulations.  
Higher angular resolution
allows GI power to spread to higher-order azimuthal modes, as demonstrated 
by a Fourier decomposition of the density.
The strengths of these modes are given by their global Fourier amplitudes,
\begin{equation}
A_m = \frac{(a_m^2 + b_m^2)^{1/2}}{\pi\int\rho_o \varpi {\rm d}\varpi {\rm d}z},
\end{equation}
where
\renewcommand{\theequation}{\arabic{equation}\alph{subequation}}
\setcounter{subequation}{2}
\begin{equation}
a_m = \int \rho \cos(m\phi)\varpi {\rm d}\varpi {\rm d}z {\rm d}\phi,
\end{equation}
\addtocounter{equation}{-1}
\addtocounter{subequation}{1}
\begin{equation}
b_m = \int \rho \sin(m\phi)\varpi {\rm d}\varpi {\rm d}z {\rm d}\phi.
\end{equation}
\renewcommand{\theequation}{\arabic{equation}}
Here  $\rho_o$ is the axisymmetric (zero-order)  component of the density.
In what follows, $A_m$ is averaged over time to 
suppress fluctuations on the dynamic time scale due
to the chaotic nature of GIs in the asymptotic  phase.
In particular, $\langle A_{tot} \rangle$ will represent 
the power over the radial range 10 -- 40 AU,
integrated from $ m = 2$ to $l_{max}/2$ and
averaged over the time range 12 -- 19 orps.  
The density distribution cannot be resolved in azimuth
on angular scales less than $2\pi/l_{max}$, therefore $A_m$ is
limited to modes $m \le l_{max}/2$.  
Because the central star is artificially fixed to the grid center, the $m = 1$ mode 
cannot be be realistically treated; it is excluded from
 $\langle A_{tot} \rangle$.

@@

The torque {\bf C} on a cylindrical surface of the disk at radius $\varpi$
can be obtained by integrating the stress tensor $T$ over the
surface of the cylinder (e.g., Lynden-Bell \& Kalnajs 1972), i.e.,
\begin{equation}
{\rm {\bf C}} = \int {\bf r} \times T \cdot {\rm d}S.
\end{equation}
If the stress tensor includes only gravitational stresses,
the surface integral can be replaced with the volume integral
\begin{equation}
{\rm \bf C} = \int \rho {\bf r} \times \nabla \Phi {\rm d}V,
\end{equation}
where $\Phi$ is the gravitational potential.  Here we are interested
only in the $z$-component of torque,  
\begin{equation}
\label{torque1}
{\rm \bf C}_Z = \int \rho \frac{\partial\Phi} {\partial \phi} {\rm d}V,
\end{equation}
as only this component  
drives mass and angular momentum transfer.  
The torque contribution from each mode $m$ can be calculated 
by replacing $\rho$ in Equation (5) with the corresponding modal density component
\begin{equation}
\rho_m = \slantfrac{1}{2} \rho_o + a_m \cos(m\phi) + b_m \sin(m\phi),
\end{equation}
where $a_m$ and $b_m$ are provided by Equation 2,
and only the gravitational potential produced by the mass distribution
given by $\rho_m$ is included in $\Phi$.
{\bf (Scott: Equation (6) is the form you use in your chapter 3, but it
does not appear to be correct; it does not yield a correct value for $m=0$)}

@@

The gravitational torque from equation \ref{torque1} can be used to determine the
gravitational stress by
\begin{equation}
\begin{split}
 T_{\varpi\phi}^{grav}  {\varpi}& = \frac{1}{2\pi\varpi^2}C_Z
\\
                                & = - \frac{1}{2\pi\varpi^2} \int \rho\frac{\partial\Phi}{\partial\phi} {\rm d}V.
\end{split}
\end{equation}
{\bf (There is a problem with this equation)}
Figure 6 shows the effective $\alpha$s for the four different grid resolutions, time
averaged from 12 to 19 orps.  
For comparison, the predicted values by Gammie (2001)
for the strongly self-gravitating limit (lower curve) and the non-self-gravitating limit
(upper curve) are shown. Note that predictions from are based upon a thin
disk approximation, where the local balance of heating and cooling dominates the
energetics.

>>>>

\usepackage{color}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{amsmath, amsthm, amssymb}

@@

Figure 2
%\ref{fig:DensityPlots}, 
shows midplane and radial densities  at $t \sim 18$ orps  
for the $l_{max} =$ 64, 128, 256, and 512 simulations. 
The lower resolution (smaller $l_{max}$) simulations
display longer wavelength spiral structures, with larger amplitudes and more coherence than those
seen in higher resolution 
simulations.  
Higher angular resolution
allows GI power to spread to higher-order azimuthal modes, as demonstrated 
by a Fourier decomposition of the density.
The strengths of these modes are given by their global Fourier amplitudes,
\begin{equation}
A_m = \frac{(a_m^2 + b_m^2)^{1/2}}{\pi\int\rho_o \varpi {\rm d}\varpi {\rm d}z},
\end{equation}
where
\renewcommand{\theequation}{\arabic{equation}\alph{subequation}}
\setcounter{subequation}{2}
\begin{equation}
a_m = \int \rho \cos(m\phi)\varpi {\rm d}\varpi {\rm d}z {\rm d}\phi,
\end{equation}
\addtocounter{equation}{-1}
\addtocounter{subequation}{1}
\begin{equation}
b_m =\int \rho \sin(m\phi)\varpi {\rm d}\varpi {\rm d}z {\rm d}\phi.
\end{equation}
\renewcommand{\theequation}{\arabic{equation}}
Here  $\rho_o$ is the axisymmetric (zero-order)  component of the density.
In what follows, $A_m$ is averaged over time to 
suppress fluctuations on the dynamic time scale due
to the chaotic nature of GIs in the asymptotic  phase.
In particular, $\langle A_{tot} \rangle$ will represent 
the power over the radial range 10 -- 40 AU,
integrated from $ m = 2$ to $l_{max}/2$ and
averaged over the time range 12 -- 19 orps.  
The density distribution cannot be resolved in azimuth
on angular scales less than $2\pi/l_{max}$, therefore $A_m$ is
limited to modes $m \le l_{max}/2$.  
Because the central star is artificially fixed to the grid center, the $m = 1$ mode 
cannot be be realistically treated; it is excluded from
 $\langle A_{tot} \rangle$.

@@

The torque {\bf C} on a cylindrical surface of the disk at radius $\varpi$
can be obtained by integrating the stress tensor $T$ over the
surface of the cylinder (e.g., Lynden-Bell \& Kalnajs 1972), i.e.,
\begin{equation}
{\rm {\bf C}} = \int {\bf r} \times T \cdot {\rm d}S.
\end{equation}
If the stress tensor includes only gravitational stresses,
the surface integral can be replaced with the volume integral
\begin{equation}
{\rm \bf C} = \int \rho {\bf r} \times \nabla \Phi {\rm d}V,
\end{equation}
where $\Phi$ is the gravitational potential.  Here we are interested
only in the $z$-component of torque,  
\begin{equation}
\label{torque1}
{\rm \bf C}_Z = \int \rho \frac{\partial\Phi} {\partial \phi} {\rm d}V,
\end{equation}
as only this component  
drives mass and angular momentum transfer.  
The torque contribution from each mode $m$ can be calculated 
by replacing $\rho$ in Equation (5) with the density distribution reconstructed from a single Fourier component
\begin{equation}
\rho_m = a_{\phi m} \cos(m\phi) + b_{\phi m} \sin(m\phi),
\end{equation}
where $a_{\phi m} =\slantfrac{1}{\pi} \int \rho \cos(m\phi){\rm d}\phi$ and $b_{\phi m} =\slantfrac{1}{\pi} \int \rho \sin(m\phi){\rm d}\phi$,
and only the gravitational potential produced by the mass distribution
given by $\rho_m$ is included in $\Phi$.
{\bf (Scott: Equation (6) is the form you use in your chapter 3, but it
does not appear to be correct; it does not yield a correct value for $m=0$)}

@@

The gravitational torque from equation \ref{torque1} can be used to determine the
gravitational stress by
\begin{equation}
\begin{split}
 T_{\varpi\phi}^{grav}\left( \varpi \right) & = \frac{1}{2\pi\varpi^2}C_Z\\
                                & = - \frac{1}{2\pi\varpi^2} \int \rho\frac{\partial\Phi}{\partial\phi} {\rm d}V.
\end{split}
\end{equation}
{\bf (There is a problem with this equation)}
Figure 6 shows the effective $\alpha$s for the four different grid resolutions, time
averaged from 12 to 19 orps.  
For comparison, the predicted values by Gammie (2001)
for the strongly self-gravitating limit (lower curve) and the non-self-gravitating limit
(upper curve) are shown. Note that predictions from are based upon a thin
disk approximation, where the local balance of heating and cooling dominates the
energetics.
